---
title: "Isotope Evaporation Line"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Put the title of your vignette here}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(devtools)
library(isoH2Obudget)
library(cslsdata)
library(dplyr) #for pipes, filtering, and selecting
library(stringr) #for str_c
library(reshape2) #for melting
library(ggplot2) #for plotting
library(extrafont) #for fonts
library(lubridate)
library(knitr)
```


```{r data, eval=TRUE, echo=FALSE}
calculate_line_intersect <- function(line1, line2){
  x <- (line1$coefficients[1] - line2$coefficients[1])/
       (line2$coefficients[2] - line1$coefficients[2])
  y <- line1$coefficients[2]*x + line1$coefficients[1]
  line_intersect <- data.frame(x = x, y = y)
  return(line_intersect)
}

dist_to_point <- function(x1, y1, x2, y2){
  dist <- sqrt((x2 - x1)^2 - (y2 - y1))
  return(dist)
}

get_inputs <- function(lake, static_gw = FALSE,
                       median_threshold = 0.01, static_lake = FALSE,
                       use_kniffin_pcpn = TRUE,
                       pcpnfile = 'csls_isotope_precipitation_deployment.csv',
                       pcpndir = "system.file") {
  
  data("weather", package = "cslsdata")
    if (lake == "Pleasant") {
      data("psnt_lst", package = "cslsdata"); lst <- psnt_lst
      data("psnt_isotopes", package = "cslsdata"); isotopes <- psnt_isotopes
      data("psnt_lake_levels", package = "cslsdata"); lake_levels <- psnt_lake_levels
      data("psnt_gw_levels", package = "cslsdata"); gw_levels <- psnt_gw_levels
      data("psnt_stage_vol", package = "cslsdata"); stage_vol <- psnt_stage_vol
      data("psnt_dictionary", package = "cslsdata"); site_dictionary <- psnt_dictionary
    } else if (lake == "Long") {
      data("long_lst", package = "cslsdata"); lst <- long_lst
      data("long_isotopes", package = "cslsdata"); isotopes <- long_isotopes
      data("long_lake_levels", package = "cslsdata"); lake_levels <- long_lake_levels
      data("long_gw_levels", package = "cslsdata"); gw_levels <- long_gw_levels
      data("long_stage_vol", package = "cslsdata"); stage_vol <- long_stage_vol
      data("long_dictionary", package = "cslsdata"); site_dictionary <- long_dictionary
    } else if (lake == "Plainfield") {
      data("pfl_lst", package = "cslsdata"); lst <- pfl_lst
      data("pfl_isotopes", package = "cslsdata"); isotopes <- pfl_isotopes
      data("pfl_lake_levels", package = "cslsdata"); lake_levels <- pfl_lake_levels
      data("pfl_gw_levels", package = "cslsdata"); gw_levels <- pfl_gw_levels
      data("pfl_stage_vol", package = "cslsdata"); stage_vol <- pfl_stage_vol
      data("pfl_dictionary", package = "cslsdata"); site_dictionary <- pfl_dictionary
    }

  # Summarise inputs over same monthly timeseries
  h2o_bal_inputs  <- summarise_inputs(weather, lst, isotopes, lake_levels,
                                      gw_levels, stage_vol, site_dictionary,
                                      static_gw, median_threshold, static_lake,
                                      use_kniffin_pcpn, pcpnfile, pcpndir)
  return(h2o_bal_inputs)
}

get_del_evap <- function(h2o_bal_inputs) {
  d18O_evap_in <- merge(monthly_weather, monthly_lst)
  d18O_evap_in <- merge(d18O_evap_in, monthly_isotopes)
  d18O_evap_in <- d18O_evap_in %>%
                  mutate(d18O_evap = calculate_d18O_evap(.data$atmp_K,
                                                         .data$ltmp_K,
                                                         .data$RH_pct,
                                                         .data$d18O_pcpn,
                                                         .data$d18O_lake)) %>%
                  select(.data$date, .data$d18O_evap)
  monthly_isotopes <- merge(monthly_isotopes, d18O_evap_in)
}

get_del_atm <- function(ltmp, del_pcpn) {
    alpha    <- d18O_evap_isotope_frac(ltmp)
    del_atm <- d18O_evap_d18O_atm(del_pcpn, alpha)
    return(del_atm)
}
```

## Overview

When looking at the **Local Meteoric Water Line** plots, keep in mind that
measurements closest to the meteoric water line are most likely inflowing
groundwater wells while measurements drifting off to the right are likely
outflowing groundwater wells (or the lake itself), which are influenced by
evaporation.

```{r schematic2, out.width = "75%", eval=TRUE, echo=FALSE, fig.align="center"}
include_graphics(system.file("images", 
                             "isotope_signatures.png", 
                             package = "isoH2Obudget", 
                             mustWork = TRUE))
```

## Lakes

### Pleasant Lake
```{r psnt, eval=TRUE, echo=FALSE}
data("site_dictionary", package = "cslsdata")
lake_ids      <- site_dictionary %>%
                 filter(.data$obs_type == "LK") %>%
                 select(.data$site_id) %>%
                 unlist()
lake_isotopes <- isotopes %>%
                 filter(.data$site_id %in% lake_ids)
pcpn_isotopes <- isotopes %>%
                 filter(.data$site_id == "PRECIP")
LMWL <- lm(d2H ~ d18O, pcpn_isotopes)
LEWL <- lm(d2H ~ d18O, lake_isotopes)

line_intersect     <- calculate_line_intersect(LMWL, LEWL)
lake_isotopes$dist <- dist_to_point(line_intersect$x, line_intersect$y,
                                    lake_isotopes$d18O, lake_isotopes$d2H)
lake_isotopes$date <- floor_date(lake_isotopes$date, unit = "month")

lake <- "Pleasant"
psnt_inputs      <- get_inputs(lake)
psnt_inputs$lake <- lake
psnt_h2o         <- calculate_h2o_bal(psnt_inputs)
psnt_h2o$lake    <- lake

lake <- "Long"
long_inputs      <- get_inputs(lake)
long_inputs$lake <- lake
long_h2o         <- calculate_h2o_bal(long_inputs)
long_h2o$lake    <- lake

lake <- "Plainfield"
pfl_inputs      <- get_inputs(lake)
pfl_inputs$lake <- lake
pfl_h2o         <- calculate_h2o_bal(pfl_inputs)
pfl_h2o$lake    <- lake

h2o    <- rbind(pfl_h2o, psnt_h2o, long_h2o)
inputs <- rbind(pfl_inputs, psnt_inputs, long_inputs)
inputs$d18O_atm <- get_del_atm(inputs$ltmp_K, inputs$d18O_pcpn)
inputs$d2H_atm <- get_del_atm(inputs$ltmp_K, inputs$d2H_pcpn)

del_evap <- lm(d2H_evap ~ d18O_evap, inputs)


lake_isotopes <- merge(lake_isotopes, h2o, by = c("date","lake"))
lake_isotopes$inputs <- lake_isotopes$P_mm + lake_isotopes$GWin
lake_isotopes$Eratio <- lake_isotopes$ET_mm/lake_isotopes$inputs


ggplot(data = lake_isotopes) +
  geom_point(aes(x = dist, y = Eratio, color = site_id))
                

ggplot() +
  geom_smooth(data = pcpn_isotopes, aes(x = d18O, y = d2H), method = "lm") +
  geom_smooth(data = lake_isotopes, aes(x = d18O, y = d2H), method = "lm") +
  geom_smooth(data = inputs, aes(x = d18O_evap, y = d2H_evap), method = "lm") +
  geom_smooth(data = inputs, aes(x = d18O_atm, y = d2H_atm), method = "lm") +
  geom_point(data = line_intersect, aes(x = x, y = y)) +
  geom_point(data = lake_isotopes, aes(x = d18O, y = d2H, color = lake)) +
  geom_point(data = inputs, aes(x = d18O_atm, y = d2H_atm, color = lake)) +
  geom_point(data = inputs, aes(x = d18O_evap, y = d2H_evap, color = lake))
 
```

## Session Info

The version of R and the versions of packages in use when this R Markdown file
was created are listed here:

```{r session_info, include=TRUE, echo=TRUE, results='markup'}
devtools::session_info()
```
