---
title: "CSLS Lake Water Budgets - Isotope Approach"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Put the title of your vignette here}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(devtools)
library(isoH2Obudget)
library(lubridate) # for dates
library(dplyr) #for %>% pipes
library(ggplot2) # for plotting
library(extrafont) # for fonts
library(reshape2) # for melting
library(NISTunits) # for unit conversions
```

```{r d18O, eval=TRUE, echo=FALSE}
data(isotopes)

# Months with data
d18O_G <- isotopes %>%
          filter(!.data$site_id %in% c("PRECIP", "CHAFFEE", "LONG", 
                                       "PLEAS", "PLAIN", "TAGATZ")) %>%
          group_by(date = floor_date(.data$date, unit = "month")) %>%
          select(date) %>%
          unique()
monthsGW <- format(d18O_G$date, "%b %Y")

d18O_L <- isotopes %>%
          filter(.data$site_id %in% c("LONG", "PLEAS", "PLAIN")) %>%
          group_by(date = floor_date(.data$date, unit = "month")) %>%
          select(date) %>%
          unique()

d18O_P <- isotopes %>%
          filter(.data$site_id == "PRECIP") %>%
          group_by(date = floor_date(.data$date, unit = "month")) %>%
          select(date) %>%
          unique()

month_start <- floor_date(min(isotopes$date), unit = "month")
month_end   <- floor_date(max(isotopes$date), unit = "month") + months(1) - days(1)
nmonths     <- get_nmonths(month_start, month_end)

have_data <- NULL
for (i in 1:nmonths){
  this_month <- month_start + months(i-1)
  have_data$Month[i] <- format(this_month, "%b %Y")
  if (this_month %in% d18O_P$date){
    have_data$Precipitation[i] <- "X"
  } else {
    have_data$Precipitation[i] <- ""
  }
  if (this_month %in% d18O_L$date){
    have_data$Lake[i] <- "X"
  } else {
    have_data$Lake[i] <- ""
  }
  if (this_month %in% d18O_G$date){
    have_data$Groundwater[i] <- "X"
  } else {
    have_data$Groundwater[i] <- ""
  }
  if (have_data$Precipitation[i] == "X" &
      have_data$Lake[i] == "X" &
      have_data$Groundwater[i] == "X"){
    have_data$All[i] <- "X"
  } else {
    have_data$All[i] <- ""
  }
}

# Precipitation - Measured
d18O_P <- isotopes %>%
          filter(.data$site_id == "PRECIP") %>%
          group_by(date = floor_date(.data$date, unit = "day")) %>%
          summarise(d18O = mean(.data$d18O),
                    d2H = mean(.data$d2H))

# Precipitation - Extended
september <- list(date = mdy("9/28/2018"), 
                  d18O = d18O_P$d18O[1], 
                  d2H = d18O_P$d2H[1])
april     <- list(date = mdy("4/19/2019"), 
                  d18O = d18O_P$d18O[3], 
                  d2H = d18O_P$d2H[3])
d18O_P_ext    <- as.data.frame(rbind(d18O_P, september, april))
d18O_P_ext    <- d18O_P[order(d18O_P$date), ]

# Precipitation - Kniffin
kniffin_isotopes <- load_pkg_csv(filename = "kniffin_isotope_samples.csv", 
                                 filedir = "system.file")
kniffin_isotopes <- kniffin_isotopes %>%
                    filter(.data$site_id == "PRECIP") %>%
                    select(.data$date, .data$d18O, .data$d2H)
kniffin_isotopes$date <- mdy(kniffin_isotopes$date)

month_abbr <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sept", 
           "Oct", "Nov", "Dec")

combo_iso <- NULL
for (i in 1:length(month_abbr)) {
  csls_pcpn           <- d18O_P_ext %>%
                         filter(month(.data$date) == i)
  if (nrow(csls_pcpn) == 0){
    csls_pcpn[1, ]    <- rep(NA, 3)
  }
  kniffin_pcpn        <- kniffin_isotopes %>%
                         filter(month(.data$date) == i)
  if (nrow(kniffin_pcpn) == 0){
    kniffin_pcpn[1, ] <- rep(NA, 3)
  }
  
  combo_iso$month[i]  <- month_abbr[i]
  combo_iso$d18O[i]   <- mean(c(csls_pcpn$d18O, kniffin_pcpn$d18O), na.rm = TRUE)
  combo_iso$d2H[i]    <- mean(c(csls_pcpn$d2H, kniffin_pcpn$d2H), na.rm = TRUE)
}

d18O_P$proj           <- "CSLS"
kniffin_isotopes$proj <- "Kniffin"

LMWL                  <- as.data.frame(rbind(kniffin_isotopes, d18O_P))

plot_LMWL <- ggplot(data = LMWL,
                    aes(x = d18O, y = d2H, group = proj)) +
             geom_point(aes(color = proj)) +
             geom_smooth(aes(color = proj),
                         method = "lm",
                         se = FALSE,
                         size = 1) + 
             scale_color_manual(name = "Data Source",
                                values = c("Kniffin" = "blue3", "CSLS" = "red3")) +
             labs(title = "Local Meteoric Water Line") +
             theme_bw() + 
             theme(text = element_text(family = "Segoe UI Semilight"))
```

## Overview

This document walks through how we're summarizing isotope data and using key
equations to calculate the water budget at each of the 3 lakes in the Central
Sands Lakes Study. Specific questions we have for you are:

1. Would you have any concerns about the way we are extending our
$\delta^{18}O_{P}$ and $\delta^{18}O_{L}$ datasets?
2. Would you suggest a different approach for defining "upgradient wells" for
$\delta^{18}O_{GW}$ measurements?
3. Would you calculate $\alpha^{*}$ using a different equation?
4. Do you have suggestions for how to adjust the theoretical $\delta^{18}O_{A}$
to account for vapour being out of equilibrium during maximum evaporation
months?
5. Are we missing any reason that calculation of $GW_{in}$ should change due to
a non-zero $dV/dt$?
6. Is it possible to calculate the monthly water balance using this approach
over the winter (ice-on) months?

## Data Sources

### Meteorological Data

Hourly precipitation and lake evaporation inputs (air temperature, relative
humidity, solar radiation, wind speed) are sourced from the [Hancock
Agricultural Research
Station](https://enviroweather.msu.edu/weather.php?stn=hck). This station is
approximately 8 miles from Plainfield Lake, 8.5 miles from Long Lake, and 14.5
miles from Pleasant Lake. Missing records for lake evaporation inputs are filled
in via linear interpolation.

### Lake Surface Temperature

Surface depth measurements from every-other-week lake temperature profiles are
averaged to yield monthly average lake surface temperature for
$\delta^{18}O_{E}$ calculations. Field measurements were not taken from November
2018 - January 2019, but we do have continuous temperature probe measurements at
0.5 m depth which we could use instead.

### Stable Isotope Measurements

**Precipitation:** $\delta^{18}O_{P}$ is measured using two collectors at the Hancock Agricultural
Research Station (approximately 8 miles from Plainfield Lake, 8.5 miles from
Long Lake, and 14.5 miles from Pleasant Lake). Collectors were initially
deployed Sept. 28, 2018 and redeployed after the winter on April 19, 2019.

**Lake:** $\delta^{18}O_{L}$ is measured less than every-other-month at the lake
surface. We do not have any profiles measurements of $\delta^{18}O_{L}$. 

**Groundwater:** $\delta^{18}O_{GW}$ is measured every-other-month at 2-4 wells that are
upgradient of each lake.

<br>

<span style="color:red">**Tricky Issue #1:</span> Months with no measurements.**
Only one month has the perfect storm of $\delta^{18}O_{P}$, $\delta^{18}O_{L}$,
and $\delta^{18}O_{GW}$ measurements.

```{r measurements, eval=TRUE, echo=FALSE, fig.height=3, fig.width=5, fig.align='center'}
knitr::kable(as.data.frame(have_data), caption = "CSLS Isotope Measurements")
```

We are extending our $\delta^{18}O_{P}$ dataset by making two assumptions:

1. For months when a) precipitation collectors were active, but b) we did not
collect a $\delta^{18}O_{P}$ sample, we assume that the measurement in the
following month also represents the month with no measurement. This extends our
coverage to include September 2018 and April 2019.
2. For all months, we average in monthly measurements of $\delta^{18}O_{P}$ made
by Maribeth Kniffin at the same location (Hancock Agricultural Resarch Station)
from May 2016 to April 2017. The Local Metoroic Water Lines from each set of
measurements match pretty well.

```{r LMWL, eval=TRUE, echo=FALSE, fig.height=3, fig.width=5, fig.align='center'}
plot_LMWL
```

<br>

We are extending our $\delta^{18}O_{L}$ dataset by making one assumption:

1. For any month with measurements of $\delta^{18}O_{P}$ and
$\delta^{18}O_{GW}$, but not $\delta^{18}O_{L}$, we are using the mean fall
$\delta^{18}O_{L}$ (i.e., the Oct 2018 measurement)

<br>

<span style="color:red">**Tricky Issue #2:</span> Which Wells are Upgradient?**
Groundwater levels have risen markedly over the last year and many wells at
Plainfield Lake and Long Lake have flipped from being "downgrdient" wells to
"upgradient" wells.

To decide if a well is upgradient or downgradient for a given
$\delta^{18}O_{GW}$ measurement, we look at the previous 30 days of daily water
level differences (lake level - groundwater level). If the median daily
difference is 1 cm or less (i.e., less than the precision of HOBO U20 loggers),
the isotope measurement is not used. Otherwise, if the mean daily difference is
greater than zero, it is considered an "upgradient" well.

For a deeper dive into how each well behaves, see
"*csls_isotope_timeseries.html*".

## Analysis

### Lake Evaporation

We are calculating lake evaporation based on the method of [McJannet et
al.](http://www.clw.csiro.au/publications/waterforahealthycountry/mdbsy/technical/U-OpenWaterEvaporation.pdf)
(2008), which [McMahon et
al.](https://www.hydrol-earth-syst-sci.net/17/1331/2013/) (2013) suggest is
appropriate for both shallow and deep lakes. CSLS lake depths range from about
2.5 m (Plainfield and Long) to about 8 m (Pleasant).

```{r ET, eval=TRUE, echo=FALSE, fig.height=4, fig.width=7, fig.align='center'}
plot_lake_ET <- function(lake){
# Data
data(weather)
data(lst)
data(isotopes)
data(water_levels)
data(site_dictionary)
data(stage_vol)

# Subset for lake
lake_subset <- subset_lake(lake, site_dictionary, lst, isotopes, 
                           water_levels, stage_vol)

# Reassign variable names
lst             <- lake_subset$lst
isotopes        <- lake_subset$isotopes
lake_levels     <- lake_subset$lake_levels
gw_levels       <- lake_subset$gw_levels
stage_vol       <- lake_subset$stage_vol
site_dictionary <- lake_subset$site_dictionary

# Monthly summary
monthly_weather   <- get_monthly_weather(weather, lst, stage_vol, 
                                         site_dictionary)

monthly_ET  <- monthly_weather %>%
               select(date = .data$date, 
                      ET = .data$ET_mm)

return(monthly_ET)
}

psnt_ET <- plot_lake_ET("Pleasant")
long_ET <- plot_lake_ET("Long")
pfl_ET  <- plot_lake_ET("Plainfield")

colnames(psnt_ET) <- c("date", "Pleasant")
colnames(long_ET) <- c("date", "Long")
colnames(pfl_ET) <- c("date", "Plainfield")

monthly_ET <- merge(psnt_ET, long_ET)
monthly_ET <- merge(monthly_ET, pfl_ET)
monthly_ET <- monthly_ET[2:nrow(monthly_ET), ]
melted_ET  <- melt(monthly_ET, id.vars = "date")

plot_ET <- ggplot(data = melted_ET, 
                  aes(x = as.Date(date), y = value, fill = variable)) +
           geom_col(position = "dodge") +
           scale_y_continuous(limits = c(0, 200),
                              expand = c(0, 0)) +
           scale_x_date(date_breaks = "2 months",
                        date_labels = "%b %y") +
           scale_fill_manual(name = "Flux",
                             values = c("steelblue4", 
                                        "steelblue3", 
                                        "steelblue2"),
                             breaks = c("Pleasant", 
                                        "Long", 
                                        "Plainfield"),
                             labels = c("Pleasant", 
                                        "Long", 
                                        "Plainfield")) +
           labs(x = "", y = "ET (mm)") +
           theme_bw() + 
           theme(text = element_text(family = "Segoe UI Semilight"))
plot_ET
```

### Groundwater Inflow

We are calculating groundwater inflow as presented in Krabbenhoft et al. (1990)
for a lake with no surface inputs.

\begin{align}
\tag{1}
GW_{in} = \frac{P*(\delta^{18}O_{L} - \delta^{18}O_{P}) + E*(\delta^{18}O_{E} - 
          \delta^{18}O_{L})}{\delta^{18}O_{GW_{in}} - \delta^{18}O_{L}}
\end{align}

where $GW_{in}$ is the groundwater inflow (mm), $P$ is precipitation (mm), $E$
is evaporation (mm), $\delta^{18}O_{L}$ is measured at the lake,
$\delta^{18}O_{P}$ is measured from precipitation, $\delta^{18}O_{GW_{in}}$ is
measured at inflowing groundwater wells, and $\delta^{18}O_{E}$ is estimated for
evaporation.

<br>

To estimate $\delta^{18}O_{E}$, the only remaining term in the equation, we are
also using the approach presented in Krabbenhoft et al. (1990).

\begin{align}
\tag{2}
\delta^{18}O_{E} = \frac{\alpha^{*} * \delta^{18}O_{L} - h*\delta^{18}O_{A} - 
                  \epsilon}{1 - h +10^{-3}\Delta \epsilon}
\end{align}

* We're calculating $h$, the relative humidity normalized to the temperature of
the surface water (-), as follows:
\begin{align}
\tag{3}
h = \frac{RH}{100}*\frac{es_{A}}{es_{L}}
\end{align}

    where $RH$ is the relative humidity (%), $es_{A}$ is the saturated vapor
pressure for the air (kPa), and $es_{L}$ is the saturated vapor pressure for the
lake (kPa). We're calculating saturated vapor pressure based on Allen et al. (1998) 
using air temperature (degrees C) or lake temperature (degrees C) for $T$ as appropriate:

\begin{align}
\tag{4}
es = 0.6108*exp(\frac{17.27T}{237 + T})
\end{align}

* We're calculating $\alpha$, the equilibrium isotope fractionation factor
($\alpha = 1/\alpha^{*}$), based on Gibson et al. (2016), where $T$ is the lake
surface temperature (K):

\begin{align}
\tag{5}
\alpha &= exp(-7.685*10^{-3} + \frac{6.7123}{T} - \frac{1666.4}{T^{2}} + \frac{350410}{T^{3}})
\end{align}

* We're calculating $\Delta \epsilon$, the kinetic fractionation factor (-), based
on Krabbenhoft et al. (1990):
\begin{align}
\tag{6}
\Delta \epsilon = 14.3*(1 - h)
\end{align}

* We're calculating $\epsilon$, the total fractionation factor (-), based on
Krabbenhoft et al. (1990) as well:

\begin{align}
\tag{7}
\epsilon = 1000(1 - \alpha^{*}) + \Delta \epsilon
\end{align}

* Lastly, we're calculating $\delta^{18}O_{A}$, the isotopic composition of the
atmosphere, based on Gibson et al. (2016):

\begin{align}
\tag{8}
\delta^{18}O_{A} &= \frac{\delta^{18}O_{P} - k*\epsilon}{1 + 10^{-3}*k*\epsilon^{+}}
\end{align}

* where $k$ can vary from 0.5 for highly seasonal climates to 1.0 for
non-seasonal climates but is set to 1.0 here and $\epsilon^{+}$ is defined as:

\begin{align}
\tag{9}
\epsilon^{+} &= (\alpha - 1)*1000
\end{align}

<br>

<span style="color:red">**Tricky Issue #3:</span> alpha equations can't match
reported values in Krabbenhoft et al. (1990).** We've checked two different
equations (both Gibson et al., 2016, Equation 16a shown above and Ozaydin et
al., 2001, Equation 9). While both agree with one another, they cannot reproduce
the values for $\alpha^{*}$ reported in Table 1 of Krabbenhoft et al. (1990)
using the lake surface temperature values also reported in Table 1. Is there a
different/more accepted way to calculate $\alpha$ or $\alpha^{*}$?

```{r fcns, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 7, fig.height = 5}
alpha_Gibson <- function(ltmp){
  alpha <- exp(-7.685e-3 + 6.7123/(ltmp) - 1666.4/(ltmp)^2 + 350410/(ltmp)^3)
  return(alpha)
}

alpha_Ozaydin <- function(ltmp){
  alpha <- exp((-2.0667 - (415.6/ltmp) + (1137/(ltmp^2))*(10^3))/1000)
  return(alpha)
}

date          <- c("4/1/86", "5/1/86", "6/1/86", "7/1/86", "8/1/86", "9/1/86", 
                   "10/1/86", "11/1/86")
ltmp          <- c(4, 11.9, 18.1, 22.2, 22.3, 18.1, 11.3, 6.0)
atmp          <- c(3.7, 11.8, 16.0, 19.4, 18.7, 12.7, 6.3, -1.3)
d18O_pcpn     <- c(-14.8, -10.7, -6.3, -5.1, -5.3, -8.4, -11.1, -13.8)
d18O_evap     <- c(-3.95, -13.90, -25.24, -23.97, -17.13, -10.40, -6.12, 33.24)
h             <- c(0.76, 0.75, 0.78, 0.84, 0.86, 0.86, 0.82, 0.90)
alphas        <- c(0.98878, 0.98956, 0.98994, 0.99024, 0.99018, 0.98964, 
                   0.98904, 0.98826)
delta_epsilon <- c(3.36, 3.36, 3.12, 2.27, 1.96, 1.96, 2.62, 1.43)
epsilon       <- c(14.68, 14.07, 13.17, 12.04, 11.78, 12.31, 13.58, 13.17)
d18O_atm      <- c(-25.88, -21.70, -17.07, -16.56, -17.46, -19.30, -22.07, -24.57)
d18O_lake      <- rep(-5.8, 8)

Krabbenhoft_data <- as.data.frame(cbind(date, d18O_evap, d18O_pcpn, d18O_atm, 
                                        d18O_lake, ltmp, atmp, h, alphas, 
                                        delta_epsilon, epsilon))
Krabbenhoft_data$date <- mdy(Krabbenhoft_data$date)

ltmp_K            <- NISTdegCtOk(ltmp)
alpha             <- alpha_Ozaydin(ltmp_K)
Ozaydin           <- as.data.frame(cbind(date, alpha))
colnames(Ozaydin) <- c("date", "alpha")
Ozaydin$date      <- mdy(Ozaydin$date)

alpha            <- alpha_Gibson(ltmp_K)
Gibson           <- as.data.frame(cbind(date, alpha))
colnames(Gibson) <- c("date", "alpha")
Gibson$date      <- mdy(Gibson$date)

cols <- c("c1" = "black", "c2" = "red3", "c3" = "blue3")

ggplot() +
  geom_point(data = Krabbenhoft_data,
             aes(x = date, 
                 y = as.numeric(as.character(alphas)),
                 color = "c1"),
             size = 3) +
  geom_point(data = Ozaydin,
             aes(x = date, 
                 y = 1/as.numeric(as.character(alpha)),
                 color = "c2"),
             size = 3) +
  geom_point(data = Gibson,
             aes(x = date, 
                 y = 1/as.numeric(as.character(alpha)),
                 color = "c3"),
             size = 3) +
  scale_color_manual(name = "Value",
                     breaks = c("c1", "c2", "c3"),
                     values = cols,
                     labels = c("Krabbenhoft", "Ozaydin", "Gibson")) +
  labs(x = "Sparkling Lake, 1986", y = "alpha*") +
  theme_bw() +
  theme(text = element_text(family = "Segoe UI Semilight"))
```

<br>

<span style="color:red">**Tricky Issue #4:</span> d18O_atm is hypothetical, not
adjusted for evaporation.** Since we did not measure $\delta^{18}O_{A}$, we must
estimate it as well. Unfortunately, the equations we're using seem to be for a
hypothetical $\delta^{18}O_{A}$ assuming vapor is in equilibrium with average
monthly precipitation at the average monthly temperature. This is a bad
assumption in Wisconsin, as is clear when we try to reproduce the values for
$\delta^{18}O_{A}$ reported in Krabbenhoft et al. (1990). Any suggestions for
how we can adjust for evaporation?

```{r d18O_atm, eval=TRUE, echo=FALSE, fig.width = 7, fig.height = 5}
d18O_atm_Gibson <- function(d18O_pcpn, alpha, k = 1) {
  epsilon_plus     <- (alpha - 1)*1000
  d18O_atm         <- (d18O_pcpn - k*epsilon_plus)/(1 + k*1e-3*epsilon_plus)
  return(d18O_atm)
}
d18O_atm         <- d18O_atm_Gibson(d18O_pcpn, 1/alphas, k = 1)
Gibson           <- as.data.frame(cbind(date, d18O_atm))
colnames(Gibson) <- c("date", "d18O_atm")
Gibson$date      <- mdy(Gibson$date)
Gibson$d18O_atm  <- as.numeric(as.character(Gibson$d18O_atm))

cols <- c("c1" = "black", "c3" = "blue3")

ggplot() +
  geom_point(data = Gibson,
             aes(x = date, 
                 y = d18O_atm,
                 color = "c3"),
             size = 3) +
  geom_point(data = Krabbenhoft_data,
             aes(x = date, 
                 y = as.numeric(as.character(d18O_atm)),
                 color = "c1"),
             size = 3) +
  scale_color_manual(name = "Value",
                     breaks = c("c1", "c3"),
                     values = cols,
                     labels = c("Krabbenhoft","Gibson")) +
  labs(x = "Sparkling Lake, 1986", y = "d18O_atm") +
  theme_bw() +
  theme(text = element_text(family = "Segoe UI Semilight"))

```

### Change in Lake Volume

<span style="color:red">**Tricky Issue #5:</span> Change in Lake Volume is not
zero.** Water levels in all three lakes have risen markedly over the past year
(>1 m in Plainfield and Long Lakes, >0.5m in Pleasant Lake), so we cannot assume
change in lake level is zero. We don't think this changes the approach for
calculating groundwater inflow, but of course we do need to account for change
in volume when calculating groundwtater outflow.

```{r lake_levels, eval=TRUE, echo=FALSE, fig.width = 7, fig.height = 5}
data("water_levels")
data("site_dictionary")
psnt <- subset_lake_levels("Pleasant", water_levels, site_dictionary)
pfl  <- subset_lake_levels("Plainfield", water_levels, site_dictionary)
long <- subset_lake_levels("Long", water_levels, site_dictionary)

psnt$lake <- "Pleasant"
pfl$lake  <- "Plainfield"
long$lake <- "Long"

lake_levels <- as.data.frame(rbind(psnt, long, pfl))

ggplot(data = lake_levels, 
       aes(x = as.Date(date), y = level_m)) +
  geom_line(color = "black") +
  facet_wrap(~lake, scales="free_y") +
  scale_x_date(date_breaks = "4 months",
               date_labels = "%b %y") +
  labs(x = "", y = "Lake Level (mamsl)") +
  theme_bw() +
  theme(text = element_text(family = "Segoe UI Semilight"))

```

### Overall Water Balance

We're planning to a) make a few updates to the lake evaporation term, and b)
convert all fluxes to volumes instead of depths, but our preliminary balances
for each lake are shown below.

<span style="color:red">**Tricky Issue #6:</span> Negative GWin over winter
months.** We're getting negative groundwater inflow values during the winter
(i.e., in February) at all lakes and we've noticed that no one seems to estimate
groundwater inflow/outflow during the winter months using this method. Right
now, we're also getting negative values for the smaller lakes (Plainfield and
Long) in October which is a little more puzzling. We're thinking we'll also have
to limit our water balance calculations to only the ice-free months, but we'd
like to get that groundwater flow component during the winter if at all
possible.

```{r psnt, eval=TRUE, echo=FALSE, fig.width = 7, fig.height = 5}
# Data
data(weather)
data(lst)
data(isotopes)
data(water_levels)
data(site_dictionary)
data(stage_vol)

# Subset for lake
lake_subset <- subset_lake(lake = "Pleasant", site_dictionary, lst, isotopes, water_levels, stage_vol)

# Reassign variable names
lst             <- lake_subset$lst
isotopes        <- lake_subset$isotopes
lake_levels     <- lake_subset$lake_levels
gw_levels       <- lake_subset$gw_levels
stage_vol       <- lake_subset$stage_vol
site_dictionary <- lake_subset$site_dictionary

# Identify monthly timeseries with complete coverage of input data
month_info      <- get_overlap_months(weather, lst, isotopes, lake_levels, gw_levels)

# Monthly summary
monthly_weather   <- get_monthly_weather(weather, lst, stage_vol, site_dictionary)
monthly_lst       <- get_monthly_lst(lst)
monthly_dV        <- get_monthly_dV(lake_levels, month_info)
monthly_isotopes  <- get_monthly_isotopes(isotopes, site_dictionary,
                                          static_wells = FALSE, lake_levels, 
                                          gw_levels, median_threshold = 0.01, 
                                          static_lake = FALSE)
monthly_isotopes  <- get_monthly_d18O_evap(monthly_weather, monthly_lst, monthly_isotopes)
monthly_h2o_bal   <- get_monthly_h2o_bal(monthly_weather, monthly_isotopes, monthly_dV, month_info)

# Plot
melted_bal <- melt(subset(monthly_h2o_bal, is.na(GWout) == FALSE), id.vars = "date")

for (i in 1:nrow(melted_bal)) {
  if (melted_bal$variable[i] == "P" |
  melted_bal$variable[i] == "GWin") {
    melted_bal$in_or_out[i] <- "In"
  } else {
    melted_bal$in_or_out[i] <- "Out"
  }
}

plot_balance <- ggplot(data = melted_bal) +
                geom_col(aes(x = in_or_out, y = value, fill = variable)) +
                facet_wrap(~date) +
                scale_y_continuous(expand = c(0,0)) +
                scale_fill_brewer(name = "Flux",
                                  palette = "Paired",
                                  breaks = c("P","E","GWin","GWout","dV"),
                                  labels = c("Precipitation",
                                             "Evapotranspiration",
                                             "Groundwater Inflow",
                                             "Groundwater Outflow",
                                             "Change in Lake Volume")) +
                labs(x = "", y = "Flux (mm)", title = "Pleasant Lake") +
                theme_bw() + 
                theme(text = element_text(family = "Segoe UI Semilight"))
  
plot_balance

```

```{r long, eval=TRUE, echo=FALSE, fig.width = 7, fig.height = 5}
# Data
data(weather)
data(lst)
data(isotopes)
data(water_levels)
data(site_dictionary)
data(stage_vol)

# Subset for lake
lake_subset <- subset_lake(lake = "Long", site_dictionary, lst, isotopes, water_levels, stage_vol)

# Reassign variable names
lst             <- lake_subset$lst
isotopes        <- lake_subset$isotopes
lake_levels     <- lake_subset$lake_levels
gw_levels       <- lake_subset$gw_levels
stage_vol       <- lake_subset$stage_vol
site_dictionary <- lake_subset$site_dictionary

# Identify monthly timeseries with complete coverage of input data
month_info      <- get_overlap_months(weather, lst, isotopes, lake_levels, gw_levels)

# Monthly summary
monthly_weather   <- get_monthly_weather(weather, lst, stage_vol, site_dictionary)
monthly_lst       <- get_monthly_lst(lst)
monthly_dV        <- get_monthly_dV(lake_levels, month_info)
monthly_isotopes  <- get_monthly_isotopes(isotopes, site_dictionary,
                                          static_wells = FALSE, lake_levels, 
                                          gw_levels, median_threshold = 0.01, 
                                          static_lake = FALSE)
monthly_isotopes  <- get_monthly_d18O_evap(monthly_weather, monthly_lst, monthly_isotopes)
monthly_h2o_bal   <- get_monthly_h2o_bal(monthly_weather, monthly_isotopes, monthly_dV, month_info)

# Plot
melted_bal <- melt(subset(monthly_h2o_bal, is.na(GWout) == FALSE), id.vars = "date")

for (i in 1:nrow(melted_bal)) {
  if (melted_bal$variable[i] == "P" |
  melted_bal$variable[i] == "GWin") {
    melted_bal$in_or_out[i] <- "In"
  } else {
    melted_bal$in_or_out[i] <- "Out"
  }
}

plot_balance <- ggplot(data = melted_bal) +
                geom_col(aes(x = in_or_out, y = value, fill = variable)) +
                facet_wrap(~date) +
                scale_y_continuous(expand = c(0,0)) +
                scale_fill_brewer(name = "Flux",
                                  palette = "Paired",
                                  breaks = c("P","E","GWin","GWout","dV"),
                                  labels = c("Precipitation",
                                             "Evapotranspiration",
                                             "Groundwater Inflow",
                                             "Groundwater Outflow",
                                             "Change in Lake Volume")) +
                labs(x = "", y = "Flux (mm)", title = "Long Lake") +
                theme_bw() + 
                theme(text = element_text(family = "Segoe UI Semilight"))
  
plot_balance

```

```{r pfl, eval=TRUE, echo=FALSE, fig.width = 7, fig.height = 5}
# Data
data(weather)
data(lst)
data(isotopes)
data(water_levels)
data(site_dictionary)
data(stage_vol)

# Subset for lake
lake_subset <- subset_lake(lake = "Plainfield", site_dictionary, lst, isotopes, water_levels, stage_vol)

# Reassign variable names
lst             <- lake_subset$lst
isotopes        <- lake_subset$isotopes
lake_levels     <- lake_subset$lake_levels
gw_levels       <- lake_subset$gw_levels
stage_vol       <- lake_subset$stage_vol
site_dictionary <- lake_subset$site_dictionary

# Identify monthly timeseries with complete coverage of input data
month_info      <- get_overlap_months(weather, lst, isotopes, lake_levels, gw_levels)

# Monthly summary
monthly_weather   <- get_monthly_weather(weather, lst, stage_vol, site_dictionary)
monthly_lst       <- get_monthly_lst(lst)
monthly_dV        <- get_monthly_dV(lake_levels, month_info)
monthly_isotopes  <- get_monthly_isotopes(isotopes, site_dictionary,
                                          static_wells = FALSE, lake_levels, 
                                          gw_levels, median_threshold = 0.01, 
                                          static_lake = FALSE)
monthly_isotopes  <- get_monthly_d18O_evap(monthly_weather, monthly_lst, monthly_isotopes)
monthly_h2o_bal   <- get_monthly_h2o_bal(monthly_weather, monthly_isotopes, monthly_dV, month_info)

# Plot
melted_bal <- melt(subset(monthly_h2o_bal, is.na(GWout) == FALSE), id.vars = "date")

for (i in 1:nrow(melted_bal)) {
  if (melted_bal$variable[i] == "P" |
  melted_bal$variable[i] == "GWin") {
    melted_bal$in_or_out[i] <- "In"
  } else {
    melted_bal$in_or_out[i] <- "Out"
  }
}

plot_balance <- ggplot(data = melted_bal) +
                geom_col(aes(x = in_or_out, y = value, fill = variable)) +
                facet_wrap(~date) +
                scale_y_continuous(expand = c(0,0)) +
                scale_fill_brewer(name = "Flux",
                                  palette = "Paired",
                                  breaks = c("P","E","GWin","GWout","dV"),
                                  labels = c("Precipitation",
                                             "Evapotranspiration",
                                             "Groundwater Inflow",
                                             "Groundwater Outflow",
                                             "Change in Lake Volume")) +
                labs(x = "", y = "Flux (mm)", title = "Plainfield Lake") +
                theme_bw() + 
                theme(text = element_text(family = "Segoe UI Semilight"))
  
plot_balance

```
