---
title: "Isotope Interpretations"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Put the title of your vignette here}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(devtools)
library(CSLSiso)
library(CSLSdata)
library(dplyr) #for pipes, filtering, and selecting
library(stringr) #for str_c
library(reshape2) #for melting
library(ggplot2) #for plotting
library(extrafont) #for fonts
library(knitr)
library(scales) # for ::
library(lubridate)

text_size <- 12
```

```{r fcns, eval=TRUE, echo=FALSE}
static_well_types <- function(lake) {
  df <- CSLSdata::dictionary[[lake]]
  df <- df %>%
        filter(obs_type == "GW") %>%
        select(site_id = site_id, 
               class = static_iso_class_detailed)
  classes <- unique(df$class)
  table   <- NULL
  for (class in classes) {
    sites <- df$site_id[df$class == class]
    sites <- str_c(as.character(sites), sep = "", collapse = ", ")
    class <- str_replace(as.character(class), "_", " ")
    class <- str_to_sentence(class)
    this_row <- c(class, sites)
    table    <- rbind(table, this_row)
  }
  table <- as.data.frame(table)
  rownames(table) <- NULL
  colnames(table) <- c("Gradient Type", "Wells")
  return(table)
}

get_well_types <- function(lake){
  df            <- runall_csls_inputs(lake)
  df            <- df %>%
                   filter(is.na(.data$GWin_sites) == FALSE &
                            .data$GWin_sites != "") %>%
                   select(.data$date, .data$GWin_sites, .data$GWout_sites)
  df$date       <- format(df$date, "%b %Y")
  colnames(df)  <- c("Month", "Upgradient Wells", "Downgradient Wells")
  return(df)
}

get_gw_gradient_d18O <- function(lake) {
  isotopes      <- CSLSdata::isotopes[[lake]]
  isotopes$date <- floor_date(isotopes$date, unit = "day")
  lake_levels   <- CSLSdata::lake_levels[[lake]]
  gw_levels     <- CSLSdata::gw_levels[[lake]]
  
  for (i in 1:nrow(gw_levels)) {
    today <- gw_levels$date[i]
    gw    <- gw_levels$level_m[i]
    lk    <- lake_levels$level_m[lake_levels$date == today]
    if (length(lk) > 0) {
      gw_levels$diff_m[i] <- gw - lk
    } else {
      gw_levels$diff_m[i] <- NA
    }
  }
  
  df <- merge(gw_levels,
              isotopes, 
              by = c("date", "site_id"), 
              all.x = TRUE, all.y = TRUE)
  df <- df %>%
        filter(!site_id %in% c("PLEAS", "LONG", "PLAIN", "PRECIP")) %>%
        select(date, site_id, diff_m, d18O)
  
  return(df)
}
```

## Overview

This document is a deeper dive into the stable isotope measurements as they
relate to one another and to the relative lake and groundwater levels.

<br>

When looking at the **water level difference** plots, keep in mind that when the
well plots above the dashed line, it is upgradient of the lake. When it plots
below the dashed line, it is downgradient of the lake. Water level measurements
are shown at a daily timestep. 

To determine if a well is upgradient or downgradient for a given measurement
record, we look at the previous month of daily water level differences. If the
median of these daily differences is less than 1cm (i.e, less than the precision
of HOBO U20 loggers), the isotope measurement is not used. Otherwise, if the
mean of the daily differences is greater than 0, it is considered an upgradient
well.

```{r schematic1, out.width = "75%", eval=TRUE, echo=FALSE, fig.align="center"}
include_graphics(system.file("images", 
                             "water_level_diff.png", 
                             package = "CSLSiso", 
                             mustWork = TRUE))
```

<br>

When looking at the **Local Meteoric Water Line** plots, keep in mind that
measurements closest to the meteoric water line are most likely inflowing
groundwater wells while measurements drifting off to the right are likely
outflowing groundwater wells (or the lake itself), which are influenced by
evaporation.

```{r schematic2, out.width = "75%", eval=TRUE, echo=FALSE, fig.align="center"}
include_graphics(system.file("images", 
                             "isotope_signatures.png", 
                             package = "CSLSiso", 
                             mustWork = TRUE))
```

## Lakes

Here we're showing all stable isotope samples for each lake broken out by site
type (precipitation, upgradient wells, downgradient wells, deep wells, and
lake).

```{r all, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4, fig.align="center"}
plot_LMWL_all(text_size = text_size)
```

### Pleasant Lake
```{r psnt, eval=TRUE, echo=FALSE}
lake = "Pleasant"

well_types       <- get_well_types(lake)
gw_gradient_d18O <- get_gw_gradient_d18O(lake)
gradient_plot    <- plot_gradients(gw_gradient_d18O, 
                                   text_size = text_size,
                                   title = sprintf("%s Lake - Water Level Comparison", lake))
LMWL_sites_plot  <- plot_LMWL_sites(lake)
```

It looks like most wells at Pleasant Lake have had a consistent relationship
with the lake over time:

```{r psnt_table, eval=TRUE, echo=FALSE}
kable(static_well_types(lake))
```

Aaron mentioned at an earlier meeting that PSNT-11 may have a well construction
issue which causes it to respond very strongly to precipitation events. We
have removed this well from analysis.

The "upgradient" wells used in the monthly water budget analysis (as determined
by relative lake levels and well levels in the 30 days prior to the measurement)
are:

```{r psnt_iso, eval=TRUE, echo=FALSE}
knitr::kable(well_types)
```

```{r psnt_plots, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=6, fig.align="center"}
gradient_plot
LMWL_sites_plot
```

### Long Lake
```{r long, eval=TRUE, echo=FALSE}
lake = "Long"
well_types       <- get_well_types(lake)
gw_gradient_d18O <- get_gw_gradient_d18O(lake)
gradient_plot    <- plot_gradients(gw_gradient_d18O, text_size = text_size,
                                   title = sprintf("%s Lake - Water Level Comparison", lake))
LMWL_sites_plot         <- plot_LMWL_sites(lake)
```

It looks like there are a few wells that are consistently upgradient of Long
Lake, but classifications are more dynamic here:

```{r long_table, eval=TRUE, echo=FALSE}
kable(static_well_types(lake))
```
 
Note that: LL-01B, LL-02B, LL-05C, and LL-09B are deeper wells drilled next to
the shallower monitoring wells. What plots as LL-05 here is called LL-05B in the
DNR water levels database.

The "upgradient" wells used in the monthly water budget analysis (as determined
by relative lake levels and well levels in the 30 days prior to the measurement)
are:

```{r long_iso, eval=TRUE, echo=FALSE}
knitr::kable(well_types)
```

```{r long_plots, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=6, fig.align="center"}
gradient_plot
LMWL_sites_plot
```


### Plainfield Lake
```{r pfl, eval=TRUE, echo=FALSE}
lake = "Plainfield"
well_types       <- get_well_types(lake)
gw_gradient_d18O <- get_gw_gradient_d18O(lake)
gradient_plot    <- plot_gradients(gw_gradient_d18O, text_size = text_size,
                                   title = sprintf("%s Lake - Water Level Comparison", lake))
LMWL_sites_plot         <- plot_LMWL_sites(lake)
```

As at Long Lake, a few wells are consistently upgradient of Plainfield Lake, but
classifications are more dynamic here:

```{r pfl_table, eval=TRUE, echo=FALSE}
kable(static_well_types(lake))
```

The "upgradient" wells used in the monthly water budget analysis (as determined
by relative lake levels and well levels in the 30 days prior to the measurement)
are:

```{r pfl_iso, eval=TRUE, echo=FALSE}
knitr::kable(well_types)
```

```{r pfl_plots, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=6, fig.align="center"}
gradient_plot
LMWL_sites_plot
```

## Session Info

The version of R and the versions of packages in use when this R Markdown file
was created are listed here:

```{r session_info, include=TRUE, echo=TRUE, results='markup'}
devtools::session_info()
```
