---
title: "Isotope Interpretations"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Put the title of your vignette here}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(devtools)
library(isoH2Obudget)
library(dplyr) #for pipes, filtering, and selecting
library(stringr) #for str_c
library(reshape2) #for melting
library(ggplot2) #for plotting
library(extrafont) #for fonts
library(knitr)
```

```{r fcns, eval=TRUE, echo=FALSE}
get_dynamic_sites <- function(lake, site_dictionary, lst, isotopes, 
                                  water_levels, stage_vol){
  lk               <- subset_lake(lake, site_dictionary, lst, isotopes, 
                                  water_levels, stage_vol)
  monthly_isotopes <- get_monthly_isotopes(lk$isotopes, lk$site_dictionary, 
                                           static_wells = FALSE, lk$lake_levels, 
                                           lk$gw_levels, median_threshold = 0.01, 
                                           static_lake = FALSE)
  monthly_isotopes            <- monthly_isotopes %>%
                                 filter(is.na(.data$GWin_sites) == FALSE &
                                          .data$GWin_sites != "") %>%
                                 select(.data$date, .data$GWin_sites)
  monthly_isotopes$date       <- as.character(monthly_isotopes$date)
  monthly_isotopes$GWin_sites <- unlist(monthly_isotopes$GWin_sites)
  upgradient_sites            <- str_c(monthly_isotopes$date, 
                                       monthly_isotopes$GWin_sites, 
                                       sep = ": ")
  return(upgradient_sites)
}
get_upgradient_sites <- function(lake, site_dictionary){
  upgradient_sites <- site_dictionary %>% 
                      filter(.data$lake == !!lake,
                             .data$static_iso_class == "upgradient") %>%
                      select(.data$static_iso_class,
                             .data$site_id)
  upgradient_sites <- str_c(unique(upgradient_sites$site_id), collapse = ", ")
  return(upgradient_sites)
}

get_water_level_diff <- function(lake, water_levels, site_dictionary){
  gw_levels   <- subset_gw_levels(lake, water_levels, site_dictionary)
  lake_levels <- subset_lake_levels(lake, water_levels, site_dictionary)
  for (i in 1:nrow(gw_levels)) {
    today <- gw_levels$date[i]
    gw    <- gw_levels$level_m[i]
    lk    <- lake_levels$level_m[lake_levels$date == today]
    if (length(lk) > 0) {
      gw_levels$diff_m[i] <- gw - lk
    } else {
      gw_levels$diff_m[i] <- NA
    }
  }
  return(gw_levels)
}

```

```{r data, eval=TRUE, echo=FALSE}
data("lst")
data("isotopes")
data("water_levels")
data("site_dictionary")
data("stage_vol")
```

## Updates
Updates since last discussion include:
 * **Timeseries of d18O**. Added a timeseries plot of d18O for each site, as in
 Krabbenhoft paper.
 * **PSNT-11.** Is now coded as "invalid" in the site dictionary so that it is
 not included in monthly isotope averages.
 * **Water level averaging rules adjusted.** Instead of looking at the
 difference between the 7 day mean lake level and 7 day mean gw level before
 each isotope measurmeent, now look at the daily difference in groundwater and
 lake levels for one month before each isotope measurement. Measurements are not
 used in monthly averages if the median difference is less than 1cm (i.e., the
 precision of HOBO U20 loggers). Otherwise, if the mean daily difference in lake
 levels and well levels is greater than zero, the measurement is classified as
 an "upgradient" measurement.

## Issues
Current issues include:

 * **Precipitation Isotopes.** Use Maribeth's values in addition to CSLS values.

## Overview

When looking at the **water level difference** plots, keep in mind that when the
well plots above the dashed line, it is upgradient of the lake. When it plots
below the dashed line, it is downgradient of the lake. Water level measurements
are shown at a daily timestep. 

To determine if a well is upgradient or downgradient for a given measurement
record, we look at the previous month of daily water level differences. If the
median of these daily differences is less than 1cm (i.e, less than the precision
of HOBO U20 loggers), the isotope measurement is not used. Otherwise, if the
mean of the daily differences is greater than 0, it is considered an upgradient
well.

```{r schematic1, out.width = "75%", eval=TRUE, echo=FALSE, fig.align="center"}
include_graphics(system.file("images", 
                             "water_level_diff.png", 
                             package = "isoH2Obudget", 
                             mustWork = TRUE))
```

<br>

When looking at the **isotope** plots, keep in mind that measurements closest to
the meteoric water line are most likely inflowing groundwater wells while
measurements drifting off to the right are likely outflowing groundwater wells
(or the lake itself), which are influenced by evaporation.

```{r schematic2, out.width = "75%", eval=TRUE, echo=FALSE, fig.align="center"}
include_graphics(system.file("images", 
                             "isotope_signatures.png", 
                             package = "isoH2Obudget", 
                             mustWork = TRUE))
```

## Pleasant Lake
```{r psnt, eval=TRUE, echo=FALSE}
lake = "Pleasant"
upgradient_sites      <- get_upgradient_sites(lake, site_dictionary)
lake_isotopes         <- subset_isotopes(lake, isotopes)
lake_isotopes$diff_m  <- 0
water_level_diff      <- get_water_level_diff(lake, water_levels, site_dictionary)
level_plot            <- plot_levels(lake, water_level_diff, lake_isotopes)
iso_plot              <- plot_iso(lake, lake_isotopes)
d18O_plot             <- plot_d18O(lake, lake_isotopes)
```

It looks like most wells at Pleasant Lake have had a consistent relationship
with the lake over time:

 * *Consistently upgradient:* PSNT-01, PSNT-14, PSNT-15, and PSNT-16
 * *Consistently downgradient:* PSNT-03, PSNT-05, PSNT-06, PSNT-07, PSNT-08,
 PSNT-09, PSNT-10
 * *Inconsistent relationship:* PSNT-11
 
Aaron mentioned at an earlier meeting that PSNT-11 may have a well construction
issue which causes it to respond very strongly to precipitation events. We
should probably remove this well from analysis.

Currently, when well classification is **static**, the only upgradient wells
used are: **`r upgradient_sites`**.

When well classification is **dynamic**, the upgradient wells each month are:
```{r psnt_iso, eval=TRUE, echo=FALSE}
get_dynamic_sites(lake, site_dictionary, lst, isotopes, water_levels, stage_vol)
```

```{r psnt_plots, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=6, fig.align="center"}
level_plot
iso_plot
d18O_plot
```

## Long Lake
```{r long, eval=TRUE, echo=FALSE}
lake = "Long"
upgradient_sites      <- get_upgradient_sites(lake, site_dictionary)
lake_isotopes         <- subset_isotopes(lake, isotopes)
lake_isotopes$diff_m  <- 0
water_level_diff      <- get_water_level_diff(lake, water_levels, site_dictionary)
level_plot            <- plot_levels(lake, water_level_diff, lake_isotopes)
iso_plot              <- plot_iso(lake, lake_isotopes)
d18O_plot             <- plot_d18O(lake, lake_isotopes)
```

It looks like there are a few wells that are consistently upgradient of Long
Lake, but classifications are more dynamic here:

 * *Consistently upgradient:* LL-01, LL-06, LL-07, LL-08
 * *Inconsistent relationship:* LL-02, LL-03, LL-04, LL-05
 * *Typically downgradient:* LL-09, LL-10
 
Note that: LL-01B, LL-02B, LL-05C, and LL-09B are deeper wells drilled next to
the shallower monitoring wells. What plots as LL-05 here is called LL-05B in the
DNR water levels database.

Currently, when well classification is **static**, the only upgradient wells
used are: **`r upgradient_sites`**.

When well classification is **dynamic**, the upgradient wells each month are:
```{r long_iso, eval=TRUE, echo=FALSE}
get_dynamic_sites(lake, site_dictionary, lst, isotopes, water_levels, stage_vol)
```

```{r long_plots, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=6, fig.align="center"}
level_plot
iso_plot
d18O_plot
```


## Plainfield Lake
```{r pfl, eval=TRUE, echo=FALSE}
lake = "Plainfield"
upgradient_sites      <- get_upgradient_sites(lake, site_dictionary)
lake_isotopes         <- subset_isotopes(lake, isotopes)
lake_isotopes$diff_m  <- 0
water_level_diff      <- get_water_level_diff(lake, water_levels, site_dictionary)
level_plot            <- plot_levels(lake, water_level_diff, lake_isotopes)
iso_plot              <- plot_iso(lake, lake_isotopes)
d18O_plot             <- plot_d18O(lake, lake_isotopes)
```

As at Long Lake, a few wells are consistently upgradient of Plainfield Lake, but
classifications are more dynamic here:

 * *Consistently upgradient:* PFL-02, PFL-04, PFL-05, PFL-15
 * *Inconsistent relationship:* PFL-07, PFL-13
 * *Typically downgradient:* PFL-03, PFL-09, PFL-11
 * *Consistently downgradient:* PFL-14

Currently, when well classification is **static**, the only upgradient wells
used are: **`r upgradient_sites`**.

When well classification is **dynamic**, the upgradient wells each month are:
```{r pfl_iso, eval=TRUE, echo=FALSE}
get_dynamic_sites(lake, site_dictionary, lst, isotopes, water_levels, stage_vol)
```

```{r pfl_plots, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=6, fig.align="center"}
level_plot
iso_plot
d18O_plot
```

## Session Info

The version of R and the versions of packages in use when this R Markdown file
was created are listed here:

```{r session_info, include=TRUE, echo=TRUE, results='markup'}
devtools::session_info()
```
