---
title: "CSLS - Exploring Isotope Measurements"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Put the title of your vignette here}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(devtools)
library(CSLSiso)
library(CSLSdata)
library(dplyr) #for pipes, filtering, and selecting
library(stringr) #for str_c
library(reshape2) #for melting
library(ggplot2) #for plotting
library(cowplot)     # for plot_grid
library(magick)      # for drawing images
library(extrafont) #for fonts
library(knitr)
library(scales) # for ::
library(lubridate)

text_size <- 12
```

```{r fcns, eval=TRUE, echo=FALSE}
static_well_types <- function(lake) {
  df <- CSLSdata::dictionary[[lake]]
  df <- df %>%
        filter(obs_type == "GW") %>%
        select(site_id = site_id, 
               class = static_iso_class_detailed)
  classes <- unique(df$class)
  table   <- NULL
  for (class in classes) {
    sites <- df$site_id[df$class == class]
    sites <- str_c(as.character(sites), sep = "", collapse = ", ")
    class <- str_replace(as.character(class), "_", " ")
    class <- str_to_sentence(class)
    this_row <- c(class, sites)
    table    <- rbind(table, this_row)
  }
  table <- as.data.frame(table)
  rownames(table) <- NULL
  colnames(table) <- c("Gradient Type", "Wells")
  return(table)
}

get_well_types <- function(lake){
  df            <- runall_csls_inputs(lake)
  df            <- df %>%
                   filter(is.na(.data$GWin_sites) == FALSE &
                            .data$GWin_sites != "") %>%
                   select(.data$date, .data$GWin_sites, .data$GWout_sites)
  df$date       <- format(df$date, "%b %Y")
  colnames(df)  <- c("Month", "Upgradient Wells", "Downgradient Wells")
  return(df)
}

get_gw_gradient_d18O <- function(lake) {
  isotopes      <- CSLSdata::isotopes[[lake]]
  isotopes$date <- floor_date(isotopes$date, unit = "day")
  lake_levels   <- CSLSdata::lake_levels[[lake]]
  gw_levels     <- CSLSdata::gw_levels[[lake]]
  
  for (i in 1:nrow(gw_levels)) {
    today <- gw_levels$date[i]
    gw    <- gw_levels$level_m[i]
    lk    <- lake_levels$level_m[lake_levels$date == today]
    if (length(lk) > 0) {
      gw_levels$diff_m[i] <- gw - lk
    } else {
      gw_levels$diff_m[i] <- NA
    }
  }
  
  df <- merge(gw_levels,
              isotopes, 
              by = c("date", "site_id"), 
              all.x = TRUE, all.y = TRUE)
  df <- df %>%
        filter(!site_id %in% c("PLEAS", "LONG", "PLAIN", "PRECIP")) %>%
        select(date, site_id, diff_m, d18O)
  
  return(df)
}
```

## Overview

This document is a deeper dive into the stable isotope measurements as they
relate to one another and to the relative lake and groundwater levels.

<br>

### Interpreting Plots

When looking at the **Water Level Difference** plots, keep in mind that when the
well plots above the dashed line, it is upgradient of the lake. When it plots
below the dashed line, it is downgradient of the lake. Water level measurements
are shown at a daily timestep. 

To determine if a well is upgradient or downgradient for a given measurement
record, we look at the previous month of daily water level differences. If the
median of these daily differences is less than 1cm (i.e, less than the precision
of HOBO U20 loggers), the isotope measurement is not used. Otherwise, if the
mean of the daily differences is greater than 0, it is considered an upgradient
well.

When looking at the **Local Meteoric Water Line** plots, keep in mind that
measurements closest to the meteoric water line are most likely inflowing
groundwater wells while measurements drifting off to the right are likely
outflowing groundwater wells (or the lake itself), which are influenced by
evaporation.

```{r schematic1, eval=TRUE, echo=FALSE, fig.align="center", fig.width=7, fig.height=4}
# include_graphics(system.file("images", 
#                              "water_level_diff.png", 
#                              package = "CSLSiso", 
#                              mustWork = TRUE))
p1 <- ggdraw() + draw_image("../inst/images/water_level_diff.png", scale = 0.92)
p2 <- ggdraw() + draw_image("../inst/images/isotope_signatures.png")
plot_grid(p1, p2, ncol = 2)
```

## Lakes

Here we're showing all stable isotope samples for each lake broken out by site
type (precipitation, upgradient wells, downgradient wells, deep wells, and
lake).

```{r all, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4, fig.align="center"}
plot_LMWL_all(text_size = text_size)
```

### Pleasant Lake

#### General Behavior
It looks like most wells at Pleasant Lake have had a consistent relationship
with the lake over time. After analyzing the graphs below, we classify each well
as follows:

```{r psnt_table01, eval=TRUE, echo=FALSE}
lake <- "Pleasant"
kable(static_well_types(lake))
```

#### Specific Wells Used Each Month
For the water budget calculations, we determine "upgradient" and "downgradient"
wells based on relative lake and well levels in the 30 days prior to the
measurement. For each month with samples, here's how we classify them:

```{r psnt_table02, eval=TRUE, echo=FALSE}
well_types <- get_well_types(lake)
knitr::kable(well_types)
```

#### Detailed Results By Well
Here's a breakdown of a) when isotope samples were collected for each well in
context with relative lake and groundwater levels, and b) how samples for each
well compare to the Local Meteoric Water Line (LMWL). Recall, we expect isotope
samples from upgradient wells to lie along the LMWL and samples from
downgradient wells to deviate from the LMWL, along the LEWL.

```{r psnt_plot01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=5.5, fig.align="center"}
gw_gradient_d18O <- get_gw_gradient_d18O(lake)
plot_gradients(gw_gradient_d18O, 
               text_size = text_size,
               title = sprintf("%s Lake - Water Level Comparison", lake))
```

```{r psnt_plot02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=5.5, fig.align="center"}
plot_LMWL_sites(lake, text_size = text_size)
```

### Long Lake

#### General Behavior
It looks like there are a few wells that are consistently upgradient of Long
Lake, but classifications are more dynamic here. After analyzing the graphs
below, we classify each well as follows:

```{r long_table01, eval=TRUE, echo=FALSE}
lake <- "Long"
kable(static_well_types(lake))
```

#### Specific Wells Used Each Month
For the water budget calculations, we determine "upgradient" and "downgradient"
wells based on relative lake and well levels in the 30 days prior to the
measurement. For each month with samples, here's how we classify them:

```{r long_table02, eval=TRUE, echo=FALSE}
well_types <- get_well_types(lake)
knitr::kable(well_types)
```

#### Detailed Results By Well
Here's a breakdown of a) when isotope samples were collected for each well in
context with relative lake and groundwater levels, and b) how samples for each
well compare to the Local Meteoric Water Line (LMWL). Recall, we expect isotope
samples from upgradient wells to lie along the LMWL and samples from
downgradient wells to deviate from the LMWL, along the LEWL.

```{r long_plot01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=6.5, fig.align="center"}
gw_gradient_d18O <- get_gw_gradient_d18O(lake)
plot_gradients(gw_gradient_d18O, 
               text_size = text_size,
               title = sprintf("%s Lake - Water Level Comparison", lake))
```

```{r long_plot02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=6.5, fig.align="center"}
plot_LMWL_sites(lake, text_size = text_size)
```

### Plainfield Lake

#### General Behavior
As at Long Lake a few wells  are consistently upgradient at Plainfield Lake, but
classifications are pretty dynamic. After analyzing the graphs below, we
classify each well as follows:

```{r pfl_table01, eval=TRUE, echo=FALSE}
lake <- "Plainfield"
kable(static_well_types(lake))
```

#### Specific Wells Used Each Month
For the water budget calculations, we determine "upgradient" and "downgradient"
wells based on relative lake and well levels in the 30 days prior to the
measurement. For each month with samples, here's how we classify them:

```{r pfl_table02, eval=TRUE, echo=FALSE}
well_types <- get_well_types(lake)
knitr::kable(well_types)
```

#### Detailed Results By Well
Here's a breakdown of a) when isotope samples were collected for each well in
context with relative lake and groundwater levels, and b) how samples for each
well compare to the Local Meteoric Water Line (LMWL). Recall, we expect isotope
samples from upgradient wells to lie along the LMWL and samples from
downgradient wells to deviate from the LMWL, along the LEWL.

```{r pfl_plot01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=5.5, fig.align="center"}
gw_gradient_d18O <- get_gw_gradient_d18O(lake)
plot_gradients(gw_gradient_d18O, 
               text_size = text_size,
               title = sprintf("%s Lake - Water Level Comparison", lake))
```

```{r pfl_plot02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=5.5, fig.align="center"}
plot_LMWL_sites(lake, text_size = text_size)
```

## Session Info

The version of R and the versions of packages in use when this R Markdown file
was created are listed here:

```{r session_info, include=TRUE, echo=TRUE, results='markup'}
devtools::session_info()
```
