---
title: "CSLS Isotope Lake Water Budget"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{CSLS_water_budget_results}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
# Libraries
library(devtools)
library(CSLSiso)
library(lubridate)
library(knitr)
library(extrafont)

# Parameters
text_size  <- 12
lakes      <- c("Pleasant", "Long", "Plainfield")
start_date <- as_datetime(mdy("09-01-2018"))
end_date   <- as_datetime(mdy("10-31-2019"))

```

## Overall Water Budget
With this, we can get the overall water budget for each lake on both a monthly
(when ice-free) and an annual basis. Groundwater is clearly a critical part of
lake water, representing over 80% of annual inflows to each lake. The volume of
flow varies quite a bit by month, with the greatest volumes of flow through
occuring during the summer (June, July, and August).

### All Lakes - Annual Balance
```{r annual, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 3}
start_date <- as_datetime(mdy("10-01-2018"))
end_date   <- as_datetime(mdy("09-30-2019"))

df <- NULL
for (lake in lakes) {
  budget      <- runall_csls_budget(lake, 
                                    annual = TRUE, 
                                    start_date = start_date,
                                    end_date = end_date)
  budget$lake <- lake
  df          <- rbind(df, budget)
}
df$lake <- factor(df$lake, levels = lakes)
plot_budget(df, 
            annual = TRUE, 
            text_size = text_size)
plot_budget(df, 
            annual = TRUE, 
            as_pcnt = TRUE, 
            text_size = text_size)

start_date <- as_datetime(mdy("09-01-2018"))
end_date   <- as_datetime(mdy("10-31-2019"))
```

### Pleasant Lake - Monthly Balance
```{r psnt01, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 5}
df <- runall_csls_budget("Pleasant", 
                         start_date = start_date, 
                         end_date = end_date)
plot_budget(df, 
            title = "Pleasant Lake",
            text_size = text_size)
```

```{r psnt02, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 3}
plot_budget(df, 
            title = "Pleasant Lake",
            text_size = text_size,
            gw_only = TRUE)
```

```{r psnt03, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 3}
plot_budget(df, 
            title = "Pleasant Lake",
            text_size = text_size,
            gw_only = TRUE,
            as_pcnt = TRUE)
```

### Long Lake - Monthly Balance
```{r long01, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 5}
df <- runall_csls_budget("Long", 
                         start_date = start_date, 
                         end_date = end_date)
plot_budget(df, 
            title = "Long Lake",
            text_size = text_size)
```

```{r long02, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 3}
plot_budget(df, 
            title = "Long Lake",
            text_size = text_size,
            gw_only = TRUE)
```

```{r long03, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 3}
plot_budget(df, 
            title = "Long Lake",
            text_size = text_size,
            gw_only = TRUE,
            as_pcnt = TRUE)
```

### Plainfield Lake - Monthly Balance
```{r pfl01, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 5}
df <- runall_csls_budget("Plainfield", 
                         start_date = start_date, 
                         end_date = end_date)
plot_budget(df, 
            title = "Plainfield Lake",
            text_size = text_size)
```

```{r pfl02, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 3}
plot_budget(df, 
            title = "Plainfield Lake",
            text_size = text_size,
            gw_only = TRUE)
```

```{r pfl03, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 3}
plot_budget(df, 
            title = "Plainfield Lake",
            text_size = text_size,
            gw_only = TRUE,
            as_pcnt = TRUE)
```

## Session Info

The version of R and the versions of packages in use when this R Markdown file
was created are listed here:

```{r session_info, include=TRUE, echo=TRUE, results='markup'}
devtools::session_info()
```
