---
title: "CSLS Isotope Lake Water Budget"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{CSLS_water_budget_isotopes}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
# Libraries
library(devtools)
library(CSLSiso)
library(CSLSdata)
library(reshape2)  # for melting
library(ggplot2)  # for plotting
library(cowplot)  # for subplots
library(extrafont)  # for fonts
library(dplyr)  # for %>% pipes
library(lubridate)
library(knitr)
library(stringr)
library(scales) # for ::

# Parameters
text_size <- 12
lakes     <- c("Pleasant", "Long", "Plainfield")

# Speed up processing by getting combinded inputs and outputs once here.
inputs  <- NULL
outputs <- NULL
for (lake in lakes){
  # Inputs
  these_inputs       <- runall_csls_inputs(lake)
  these_inputs$lake  <- lake
  inputs             <- rbind(inputs, these_inputs)
  # Outputs
  these_outputs      <- runall_csls_budget(lake)
  these_outputs$lake <- lake
  outputs            <- rbind(outputs, these_outputs)
}
inputs$lake  <- factor(inputs$lake, levels = lakes)
outputs$lake <- factor(outputs$lake, levels = lakes)
```

## Overview {#top}
This **water budget** model uses 1) observed changes in lake levels, 2) observed
precipitation, 3) calculated lake evaporation, and 4) stable isotope
measurements of precipitation, evaporation, lake water, and inflowing
groundwater to calculate monthly groundwater flow into and out of a lake.

<br>

The lakes in the Central Sands Lakes Study (Pleasant Lake, Long Lake, and
Plainfield Lake) are all seepage lakes, a type of lake that does not have
inflowing streams, outflowing streams, or appreciable surface runoff. The
water budget for a seepage lake is as follows:

\begin{align}
\tag{1}
\frac{dV}{dt} = P - E + GW_{in} - GW_{out}
\end{align}

where $\frac{dV}{dt}$ is the change in lake volume (m^3^), $P$ is the
precipitation (m^3^), $E$ is the lake evaporation (m^3^), $GW_{in}$ is the
groundwater flowing in to the lake (m^3^), and $GW_{out}$ is the groundwater
flowing out of the lake.

<br>

[**Change in Lake Volume**](#dV) is *directly measured* using a water level
logger at each lake. Daily lake levels elevations are converted into daily lake
volumes using lake elevation-volume relationships developed from lake bathymetry
maps.

<br>

[**Precipitation**](#precip) is *directly measured* at a nearby weather station.
It is converted to a volume by multiplying the total precipitation (mm) by the
average lake area during the month of interest. Average lake area is determined
using daily lake level elevation from water level loggers and the lake
elevation-area relationships developed from bathymetry maps.

<br>

[**Lake Evaporation**](#evap) is *calculated* based on the lake energy-budget
approach outlined in McJannet et al. (2008) as described in McMahon et al.
(2013). This approach uses:

* Measured incoming solar radiation, wind speed, relative humidity, and air
temperature from a nearby weather station
* Lake surface temperature from continuous temperature loggers
* Area and depth of the lake from daily lake stage measurements and
elevation-area relationships

Evaporation as a depth (Eq. 2) is summarized at a monthly timestep and converted
to a volume using the average lake area during that month, as in the
precipitation calculations.

\begin{align}
\tag{2}
E = \frac{\Delta_{w}*(R_{n} - G_{w}) + 
          \frac{60*60*24*\rho_{a}*c_{a}*(es_{w} - es_{a})}{r_{a}}}{\lambda*(\Delta_{w} + \gamma)}
\end{align}

where $E$ is lake evaporation (mm/day), $\Delta_{w}$ is the slope of the
saturation water vapour curve at water temperature (kPa/deg C), $R_{n}$ is the
net radiation (MJ/m^2^/day), $G_{w}$ is the change in heat storage in the water
body (MJ/m^2/day), $\rho_{a}$ is the density of air (kg/m^3^), $c_{a}$ is the
specific heat of air (MJ/kg/K), $es_{w}$ is the saturation vapour pressure at
water temperature (kPa), $es_{a}$ is the saturation vapour pressure of the air
(kPa), $r_{a}$ is the aerodynamic resistence (s/m), $\lambda$ is the latent heat
of vapourisation (MJ/kg), and $\gamma$ is the psychormetric constant (kPa/deg
C).

<br>

[**Groundwater inflow**](#GWin) is *calculated* using monthly changes in lake
volume, precipitation, lake evaporation, and stable isotope measurements based
on the approach in Krabbenhoft et al. (1990) and modified based on Mook (2000) to
account for non-steady state conditions in lake levels:

\begin{align}
\tag{3}
GW_{in} = \frac{P*(\delta^{18}O_{L} - \delta^{18}O_{P}) + E*(\delta^{18}O_{E} - 
          \delta^{18}O_{L}) + V*d\delta^{18}O_{L}/dt}{\delta^{18}O_{GW_{in}} - \delta^{18}O_{L}}
\end{align}

where $GW_{in}$ is the groundwater inflow (m^3^), $P$ is precipitation (m^3^), $E$
is evaporation (m^3^), V is the lake volume (m^3^) and $\delta^{18}O_{x}$ is the
relative isotopic composition in units of per mil relative to a known standard.
$\delta^{18}O_{L}$ is measured at the lake, $\delta^{18}O_{P}$ is measured from
precipitation, $\delta^{18}O_{E}$ is estimated for evaporation, and
$\delta^{18}O_{GW_{in}}$ is measured at inflowing groundwater wells.

<br>

[**Groundwater outflow**](#GWout) can then be *calculated* as the only remaining
term in the water balance:

\begin{align}
\tag{4}
GW_{out} = GW_{in} + P - E - \frac{dV}{dt}
\end{align}

## Detailed Approach
Input data for the Central Sands Lake Study sites (Pleasant Lake, Long Lake, and
Plainfield Lake) are saved in the `CSLSdata` R data package. The `CSLSdata`
package includes a `data-raw` directory where raw data is stored together with
the scripts used to clean raw data and save them as Rda files. Detailed
information on the source of each dataset is captured in the `CSLSdata` manual.
This is accessible as a pdf or via RStudio (`help(package = CSLSdata)`).

<br>

Below, we describe all additional assumptions and calculations used to transform
`CSLSdata` Rda files from their form in the `CSLSdata` package into the monthly
inputs required for monthly water balance calculations.

### Change in Lake Volume {#dV}
[back to top](#top)

We obtain **daily lake elevation data** from water level loggers at USGS gages
at all three lakes in the Central Sands Lakes Study.

```{r dV01, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 2.5}
plot_lake_level(text_size = text_size)
```

We also have the **elevation-volume relationships** for each lake from lake
bathymetry data (calculated in ArcGIS).

```{r dV02, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 2.5}
plot_elev_area_vol(ydata = "vol_m3", 
                   ylabel = expression(Lake~Volume~(m^{3})), 
                   text_size = text_size)
```

We convert these daily lake level elevations to **daily lake volumes** using these elevation-volume relationships.

```{r dV03, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 2.5}
plot_lake_level(ydata = "vol_m3",
                ylabel = expression(Lake~Volume~(m^{3})), 
                yscales = scales::scientific, 
                text_size = text_size)
```

The **change in lake volume** for a given month is then the difference
between the lake volume on the last day of the month and the lake volume on the
first day of the month.

```{r dV04, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 2.5}
plot_input(df = inputs,
           ydata = "dV_m3", 
           ylabel = expression(Change~Lake~Volume~(m^{3})), 
           geom_type = "col", 
           text_size = text_size)
```

### Precipitation {#precip}
[back to top](#top)

We obtain hourly precipitation depth (mm) from the [Hancock Agricultural
Research Station](https://enviroweather.msu.edu/weather.php?stn=hck) (station
id: hck). The station is located in Hancock, WI (location: 44.1188, -89.533,
elevation: 241m), approximately 8 miles from Plainfield Lake, 8.5 miles from
Long Lake, and 14.5 miles from Pleasant Lake.

```{r precip0, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 2.5}
weather <- CSLSdata::weather
monthly_weather <- weather %>%
                   group_by(date = floor_date(.data$date, unit = "month")) %>%
                   summarise(P = sum(.data$P))
ggplot(monthly_weather, aes(x = date, y = P)) +
  geom_col() +
  scale_x_datetime(date_breaks = "3 months",
                   date_labels = "%b '%y") +
  scale_y_continuous(labels = scales::scientific) +
  labs(x = "",
       y = "Monthly Precipitation (mm)") +
  theme_bw() +
  theme(text = element_text(family = "Segoe UI Semilight",
                            size = text_size),
        plot.title = element_text(hjust = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```

<br>

To convert precipitation as a depth to precipitation as a volume, We use the
**daily lake elevations** at the USGS gages at each lake.

```{r precip1, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 2.5}
plot_lake_level(text_size = text_size)
```

We also use the **elevation-area relationships** for each lake from lake
bathymetry data (calculated in ArcGIS).

```{r precip2, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 2.5}
plot_elev_area_vol(ydata = "area_m2", 
                   ylabel = expression(Lake~Area~(m^{2})), 
                   text_size = text_size)
```

We convert daily lake level elevations to **daily lake areas** using these
elevation-area relationships.

```{r precip3, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 2.5}
plot_lake_level(ydata = "area_m2", 
                ylabel = expression(Lake~Area~(m^{2})), 
                yscales = scales::scientific, 
                text_size = text_size)
```

Then, we calculate the **mean monthly lake area**.

```{r precip4, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 2.5}
plot_input(df = inputs,
           ydata = "mean_area_m2", 
           ylabel = expression(Lake~Area~(m^{2})), 
           geom_type = "point", 
           facet_scales = "free_y",
           text_size = text_size)
```

By multiplying monthly precipitation as a depth (mm) by the mean monthly lake
area (m^2^), we obtain **monthly precipitation as a volume (m^3^)** for each lake.

```{r precip5, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 2.5}
plot_input(df = inputs,
           ydata = "P_m3", 
           ylabel = expression(Precipitation~(m^{3})), 
           geom_type = "col", 
           text_size = text_size)
```

### Lake Evaporation {#evap}
[back to top](#top)

All lake evaporation calculations are contained in the `CSLSevap` package; more
details on this energy-balance approach can be obtained by searching the
documentation for that package. Broadly, the energy-budget lake evaporation
equations require information on:

* Lake location (latitude, longitude, elevation, latitude of time zone)
* Lake size (area, depth)
* Lake surface temperature (daily values, taken from surface-most continuous
temperature logger)
* Meteorological parameters (daily relative humidity, air temperature, incoming solar
radiation, wind speed from nearby weather station)
* Heat absorption and reflection parameters (e.g., albedo of water, density and
specific heat of air)

These lake evaporation equations yeild daily lake evaporation (mm/day), which is
converted to monthly lake evaporation volume by summing over each month and
multiplying by the average monthly lake area, as in the [precipitation](#precip)
calculations.

```{r evap, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 2.5}
plot_input(df = inputs,
           ydata = "E_m3", 
           ylabel = expression(Evaporation~(m^{3})), 
           geom_type = "col", 
           text_size = text_size)
```

### Groundwater Inflow {#GWin}
[back to top](#top)

As indicated in Eq. 3, calculation of groundwater inflow requires:

1. Monthly **precipitation**, **evaporation**, and **mean lake volume**. These
are calculated per the previous sections.

2. Monthly **groundwater stable isotope measurements** are derived from regular
CSLS sampling at four groundwater wells near each lake. These measurements were
taken approximately every other month. Upgradient wells representing inflowing
groundwater are defined as wells where the mean daily difference between
groundwater levels and lake levels in the 30-day period prior to sampling is
greater than 1 cm (i.e., the precision of the water level logger). All
upgradient well measurements from the same sampling event for a lake are
averaged, then linearly interpolated to estimate values on the last day of each
month.

3. Monthly **lake stable isotope measurements** are derived from regular CSLS
sampling at each lake. These measurements were taken less frequently than
groundwater measurements; approximately every other month during the ice-free
period. All lake measurements are linearly interpolated to estimate values on
the last day of each month during the ice-free period.

4. Monthly **precipitation stable isotope measurements** are derived from
regular CSLS sampling at two precipitation collectors located at Hancock
Agricultural Research Station. These measurements were taken approximately every
month during the snow-free season. They are supplemented by monthly measurements
taken by Maribeth Kniffin at the same location from May 2016 to April 2017.
Since 1) we do not have access to the exact days that Maribeth Kniffin collected
her measurements and 2) precipitation samples are an aggregate sample of all
precipitation collected between sample dates (i.e., they are not a snapshot in
time), we do not perform any linear interpolation on these samples. Instead, we
assume all precipitation measurements represent precipitation on the last day of
the month in which they were collected. Maribeth Kniffin's data is only used for
months when we do not have CSLS-collected data.

```{r iso, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 7.5}
lakes <- c("Pleasant", "Long", "Plainfield")
monthly_isotopes <- NULL
for (lake in lakes){
  isotopes    <- CSLSdata::isotopes[[lake]]
  dictionary  <- CSLSdata::dictionary[[lake]]
  lake_levels <- CSLSdata::lake_levels[[lake]]
  gw_levels   <- CSLSdata::gw_levels[[lake]]
  
  # Measured values
  tmp1 <- iso_site_type(isotopes, dictionary, lake_levels, gw_levels)
  tmp1 <- tmp1 %>%
          group_by(floor_date(.data$date, unit = "month"),
                   variable = .data$site_type) %>%
          filter(is.na(.data$site_type) == FALSE,
                 .data$site_type != "",
                 .data$site_type != "downgradient",
                 .data$site_type != "deep") %>%
          summarise(date = floor_date(mean(.data$date), unit = "day"),
                    value = mean(.data$d18O)) %>%
          ungroup() %>%
          select(.data$date, .data$variable, .data$value)
  
  tmp1$variable <- tmp1$variable %>%
                   str_replace("upgradient","Groundwater") %>%
                   str_replace("lake","Lake") %>%
                   str_replace("precipitation","Precipitation")
  tmp1$lake <- lake
  tmp1$measurement <- "actual"
  
  # monthly values
  tmp2 <- summarise_isotopes(isotopes, dictionary, 
                             find_timeseries(isotopes), 
                             lake_levels, gw_levels) %>% 
          select(date, d18O_GWin, d18O_lake, d18O_pcpn) %>% 
          melt(id.vars = "date")
  
  tmp2$variable    <- tmp2$variable %>%
                      str_replace("d18O_GWin","Groundwater") %>%
                      str_replace("d18O_lake","Lake") %>%
                      str_replace("d18O_pcpn","Precipitation")
  tmp2$lake        <- lake
  tmp2$measurement <- "monthly"
  
  tmp3      <- tmp1 %>% filter(variable == "Precipitation")
  tmp3      <- fill_timeseries_gaps(tmp3, find_timeseries(isotopes))
  kniffin_isotopes <- CSLSdata::kniffin$isotopes
  for (i in 1:nrow(tmp3)) {
    this_month    <- month(tmp3$date[i])
    this_i        <- which(month(tmp3$date) == this_month)
    tmp3$value[i]    <- mean(tmp3$value[this_i], na.rm = TRUE)
    tmp3$measurement[i] <- "actual"
    tmp3$value[is.nan(tmp3$value)] <- NA
    this_kniffin <- kniffin_isotopes %>% 
                     filter(month(.data$date) == this_month,
                            .data$site_id == "PRECIP")
    if (nrow(this_kniffin) == 0) {
      kniffin_d18O <- NA
    } else {
      kniffin_d18O <- mean(this_kniffin$d18O, na.rm = TRUE)
    }
    if (is.na(tmp3$value[i])){
      tmp3$value[i]       <- kniffin_d18O
      tmp3$measurement[i] <- "kniffin"
    }
  }
  tmp3$lake <- lake
  tmp3$variable <- "Precipitation"
  
  monthly_isotopes <- rbind(monthly_isotopes, tmp1, tmp2, tmp3)
}

monthly_isotopes$lake <- factor(monthly_isotopes$lake, levels = lakes)

ggplot(monthly_isotopes %>% arrange(measurement), 
       aes(x = date, y = value, color = measurement, shape = measurement)) +
  geom_point() +
  scale_x_datetime(date_breaks = "4 months",
                   date_labels = "%b '%y") +
  facet_grid(variable~lake) +
  labs(x = "", y = expression(delta^18*'O'), title = "") +
  scale_color_manual(name = "",
                     values = c("red", "purple", "grey70"),
                     labels = c("Measured by CSLS", 
                                "Measured by M. Kniffin", 
                                "Monthly Interpolation")) +
  scale_shape_manual(name = "",
                     values = c(19, 15, 17),
                     labels = c("Measured by CSLS", 
                                "Measured by M. Kniffin", 
                                "Monthly Interpolation")) +
  theme_bw() +
  theme(text = element_text(family = "Segoe UI Semilight",
                            size = text_size),
        plot.title = element_text(hjust = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "top")
  
```

5. Monthly **evaporation stable isotope measurements** must be estimated since
it cannot be directly measured. We estimate $\delta^{18}O_{E}$ as follows
(Krabbenhoft et al., 1990, Eq. 5):

   \begin{align}
   \tag{5}
   \delta^{18}O_{E} = \frac{(1/\alpha) * \delta^{18}O_{L} - h*\delta^{18}O_{A} - 
                      \epsilon}{1 - h +10^{-3}\Delta \epsilon}
   \end{align}

   The relative humidity normalized to the temperature of the surface water (-),
   $h$, is calculated as follows (Mook, 2000, Eq. 1.8):
   \begin{align}
   \tag{6}
   h = \frac{RH}{100}*\frac{es_{A}}{es_{L}}
   \end{align}

   where $RH$ is the relative humidity (%), $es_{A}$ is the saturated vapor
   pressure for the air (kPa), and $es_{L}$ is the saturated vapor pressure for
   the lake (kPa). Saturated vapor pressure is calculated based on Allen et al. (1998) 
   using air temperature (degrees C) or lake temperature (degrees C) for $T$ as 
   appropriate:
   \begin{align} 
   \tag{7} 
   es = 0.6108*exp(\frac{17.27T}{237 + T})
   \end{align}

   The equilibrium isotope fractionation factor (-), $\alpha$, is calculated
   based on Gibson et al. (2016), where $T$ is the lake surface temperature
   (K). Note that this formulation represents the fraction of isotopes in liquid 
   vs. the fraction in vapor (i.e., L/V form, $\alpha > 1$). Others, including 
   Krabbenhoft et al. (1990), use the inverse form of $\alpha$ (i.e., V/L form, 
   $\alpha < 1$) in their equations. In all equations presented in this 
   document, we consistently use the L/V form for alpha. If you calculate the 
   V/L form of alpha, simply replace $\alpha$ with $\frac{1}{\alpha}$ in 
   the equations presented in this document.

   \begin{align}
   \tag{8}
   \alpha &= exp(-7.685*10^{-3} + \frac{6.7123}{T} - \frac{1666.4}{T^{2}} + \frac{350410}{T^{3}})
   \end{align}

   The kinetic fractionation factor (-), $\Delta \epsilon$, is calculated based
   on Krabbenhoft et al. (1990):
   \begin{align}
   \tag{9}
   \Delta \epsilon = 14.3*(1 - h)
   \end{align}

   The total fractionation factor (-), $\epsilon$, combines both the isotopic
   fractionation factor and the kinetic fractionation factor (Krabbenhoft et
   al., 1990):
   \begin{align}
   \tag{10}
   \epsilon = 1000(1 - \frac{1}{\alpha}) + \Delta \epsilon
   \end{align}

   Lastly, the isotopic composition of the atmosphere, $\delta^{18}O_{A}$, is
   calculated based on Gibson et al. (2016):
   \begin{align}
   \tag{11}
   \delta^{18}O_{A} &= \frac{\delta^{18}O_{P} - k*\epsilon^{+}}{1 + 10^{-3}*k*\epsilon^{+}}
   \end{align}

   where $k$ can vary from 0.5 for highly seasonal climates to 1.0 for
   non-seasonal climates but is set to 1.0 here and $\epsilon^{+}$ is defined
   as:
   \begin{align}
   \tag{12}
   \epsilon^{+} &= (\alpha - 1)*1000
   \end{align}

```{r iso_evap, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 2.5}
plot_input(df = inputs,
           ydata = "d18O_evap", 
           ylabel = expression(delta^18*'O'[evap]), 
           yscales = scales::number,
           geom_type = "point", 
           text_size = text_size)
```

<br>

With the above information, we can then calculate groundwater inflow using Eq. 3. 

```{r GWin, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 2.5}
plot_input(df = outputs,
           ydata = "GWin_m3", 
           ylabel = expression(Groundwater~Inflow~(m^{3})), 
           inputs_only = FALSE, 
           text_size = text_size)
```

### Groundwater Outflow {#GWout}
[back to top](#top)

Finally, we calculate groundwater outflow as the only remaining term of the
water balance per Eq. 4.

```{r GWout, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 2.5}
plot_input(df = outputs,
           ydata = "GWout_m3", 
           ylabel = expression(Groundwater~Outflow~(m^{3})), 
           inputs_only = FALSE, 
           text_size = text_size)
```

## References

Allen, R. G., Pereira, L. S., Raes, D., & Smith, M. (1998). Crop
evapotranspiration: Guidelines for computing crop water requirements. Rome: FAO.
Retrieved from http://www.fao.org/docrep/X0490E/x0490e00.htm.

Gibson, J.J., S.J. Birks, and Y. Yi. 2016. Stable isotope mass balance of lakes:
a contemporary perspective. Quaternary Science Reviews, 131:316-328.
https://doi.org/10.1016/j.quascirev.2015.04.013

Krabbenhoft, D. P., C. J. Bowser, M. P. Anderson, and J. W. Valley. (1990).
Estimating Groundwater Exchange with Lakes: 1. The Stable Isotope Mass Balance
Method. Water Resources Research, 26(10):2445-2453.
https://doi.org/10.1029/WR026i010p02445

McJannet, D. L., Webster, I. T., Stenson, M. P., and Sherman, B.S. (2008).
Estimating open water evaporation for the Murray-Darling Basin. A report to the
Australian Government from the CSIRO Murray-Darling Basin Sustainable Yields
Project, CSIRO, Australia, 50 pp. Retrieved from
http://www.clw.csiro.au/publications/waterforahealthycountry/mdbsy/technical/U-OpenWaterEvaporation.pdf.

McMahon, T. A., Peel, M. C., Lowe, L., Srikanthan, R., and McVicar, T. R.
(2013). Estimating actual, potential, reference crop and pan evaporation using
standard meteorological data: a pragmatic synthesis, Hydrol. Earth Syst. Sci.,
17, 1331–1363. https://doi.org/10.5194/hess-17-1331-2013.

Mook, W.G. (ed.) 2000. Environmental Isotopes in the Hydrologic Cycle: Volume
III: Surface Water. UNESCO. Paris, France.

## Session Info

The version of R and the versions of packages in use when this R Markdown file
was created are listed here:

```{r session_info, include=TRUE, echo=TRUE, results='markup'}
devtools::session_info()
```
